Article Title

### 2014-01-27 plos.tex
Kreso's edits

\section*{Discussion}
\paragraph{Functional connectivity as a network of pairwise interactions}
Functional connectivity is often represented as a graph of pairwise interactions. The goal of many studies of functional connectivity was to estimate  anatomical connectivity from  observed multineuronal spiking activity.  For example, characteristic peaks and troughs in the pairwise cross-correlograms of recorded spike trains contain statistical signatures of directional monosynaptic connections and shared synaptic inputs \cite{Gerstein:1964, Perkel:1967, Moore:1970, Alonso:1998, Denman:2013}.  Such signatures are ambiguous as they can arise from network effects other than direct synaptic connections \cite{Aertsen:1989}.  With simultaneous recordings from more neurons, ambiguities can be resolved by inferring the conditional dependencies between pairs of neurons.  Direct causal interactions between neurons produce statistical dependency between them even after conditioning on the state of the remainder of the network. Therefore, conditionally independent neurons can be inferred to lack a direct causal influence.

Conditional dependencies can be inferred by fitting a probabilistic model of the joint population activity. For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration and membrane kinetics, and individual neurons' stimulus drive~\cite{Pillow:2008}.  Maximum entropy models are similarly constrained by observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009} .  
Under the multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.   Each probabilistic model, fitted to the same data may reveal a different network of \sq{interactions},  \ie conditional dependencies between pairs of cells.

It is not yet clear which approach provides the best correspondence with anatomical connectivity. Little experimental evidence is available to answer this question.  The connectivity graphs inferred by various statistical methods are commonly reported as abstract representations of the functional connectivity without examining their relation to anatomical implications.  
Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.  However, the topological properties of functional connectivity graphs vary with the method of their inference \cite{Zalesky:2012}. Until a physiological interpretation of functional connectivity is established, the validity of such analyses remains in question.

Inference of the conditional dependencies also depends on the completeness of the recorded population:  To properly condition the pairwise interactions on the activity of the other neurons, we must record from all neurons with which the pair interacts. Unobserved portions of the circuit can introduce false conditional dependencies between observed neurons. For this reason, the most statistical models of population activity have been most successfully applied to \emph{in vitro} preparations of the retina or cell cultures where high-quality recordings from the complete populations were available \cite{Pillow:2008}. In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of the inference of the pairwise conditional dependencies. This may be the reason that partial correlations have not previously been used to describe  functional in cortical populations.

Two-photon imaging of population calcium signals presents unique advantages for the estimation of functional connectivity.  While the temporal resolution of calcium signals is limited by calcium dye kinetics, fast imaging techniques combined with spike inference algorithms provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010}. However, such high temporal precision comes at the cost of the accuracy of inferred spike rates.  Better accuracy is achieved when calcium signals are analyzed on scales of tens of milliseconds \cite{Cotton:2013}.  The major advantage of calcium imaging is its ability to characterize the spatial arrangement and types of recorded cells.  Recently, advanced imaging techniques have allowed for recordings from nearly every cell in a volume of cortical tissue in vivo \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.  These techniques may provide more incisive measurements of functional connectivity than classical electrophysiological recordings.

@@

Coactivation patterns and pairwise connectivity are not mutually exclusive since assemblies arise from patterns of synaptic connectivity, but an analysis of coactivation shifts the focus from detailed interactions to  collective behavior.  
In our study, the analysis of functional connectivity through modes of coactivations was represented by the factor analysis estimator $C_{\sf factor}$.

@@

\paragraph{Addressing ill-posedness}
Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data produce large errors in the inferred network of dependencies.  The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.  As  techniques improved to allow for recording from larger neuronal populations, experimental neuroscientists addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted and that the stationarity assumption is observed\cite{Tkacik:2013}.  However,  ambitious projects such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations. Simply increasing recording duration will not be sufficient, and the problem must be addressed using regularized estimators. Regularization biases the solution toward a smaller subspace to counteract the effect of the sampling noise in the empirical data. However, biasing the solution to an inappropriate subspace does not allow significant estimation improvement and hampers interpretation.

@@

The results of model selection must be interpreted with caution.  As we demonstrated by simulation, a wrong simple model can produce a better estimate than a more complex model with correct representation of dependencies.   Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not support the conclusion that it reveals a better representation of dependencies in the data.  This caveat is related to \emph{Stein's Paradox} \cite{Efron:1977}: in which the biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement over the best unbiased estimate.

@@

Moreover, this estimation approach provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit. Importantly, the inferred functional interactions were substantially different from those measured by considering the network of the most significant sample correlations \figref{05}{F, H, and I}.  For example, the $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that were not present in the sample correlation matrix (\figref{05}{F}) that may reflect inhibitory circuitry.  Although the network of functional interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, will not be identical to the anatomical connectivity it may accurately reflect average connectivity profiles. Future experiments utilizing molecular markers for different cell types, and follow-up multiple whole-cell in vitro experiments \cite{Hofer:2012, Ko:2013} will directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry. Finally, the latent units inferred by the estimator can be analyzed for their physiological interpretation. For example, these latent units may be modulated under different brain states (e.g. slow-wave sleep, attention) and stimulus conditions (e.g. certain types of stimuli may engage feedback connections)

>>>>

\section*{Discussion}
\paragraph{Functional connectivity as a network of pairwise interactions}
Functional connectivity is often represented as a graph of pairwise interactions. The goal of many studies of functional connectivity was to estimate  anatomical connectivity from  observed multineuronal spiking activity.  For example, characteristic peaks and troughs in the pairwise cross-correlograms of recorded spike trains contain statistical signatures of directional monosynaptic connections and shared synaptic inputs \cite{Gerstein:1964, Perkel:1967, Moore:1970, Alonso:1998, Denman:2013}.  Such signatures are ambiguous as they can arise from network effects other than direct synaptic connections \cite{Aertsen:1989}.  With simultaneous recordings from more neurons, ambiguities can be resolved by inferring the conditional dependencies between pairs of neurons.  Direct causal interactions between neurons produce statistical dependency between them even after conditioning on the state of the remainder of the network \Kcomment{and the input?}. Therefore, conditionally independent neurons can be inferred to lack a direct causal influence.

Conditional dependencies can be inferred by fitting a probabilistic model of the joint population activity. For example, generalized linear models (GLMs) have been constructed to  include biophysically plausible synaptic integration and membrane kinetics, and individual neurons' stimulus drive~\cite{Pillow:2008}.  Maximum entropy models are similarly constrained by \Kcomment{stimulus drive and} observed pairwise correlations \cite{Schneidman:2006, Tkacik:2006, Yu:2008, Tang:2008, Shlens:2009}.  
\Kcomment{Assuming that the population response follows a} multivariate normal distribution, the conditional dependencies between pairs of neurons are expressed by the partial correlations between them.   Each probabilistic model, fitted to the same data may reveal a different network of \sq{interactions},  \ie conditional dependencies between pairs of cells.

It is not yet clear which approach provides the best correspondence with anatomical connectivity. Little experimental evidence is available to answer this question.  The connectivity graphs inferred by various statistical methods are commonly reported without examining their relation to \Kcomment{anatomy}.  
Topological properties of such graphs have been interpreted as principles of circuit organization (\emph{e.g.} small-world organization) \cite{Feldt:2011, Yu:2008, Malmersjo:2013}.  However, the topological properties of functional connectivity graphs can depend on the method of inference \cite{Zalesky:2012}. Until a physiological interpretation of functional connectivity is established, the validity of such analyses remains in question.

Inference of the conditional dependencies also depends on the completeness of the recorded population:  To properly condition the pairwise interactions on the activity of the other neurons, we must record from all neurons with which the pair interacts. Unobserved portions of the circuit can introduce false conditional dependencies between observed neurons. For this reason, statistical models of population activity have been most successfully applied to \emph{in vitro} preparations of the retina or cell cultures where high-quality recordings from the complete populations were available \cite{Pillow:2008}. In cortical tissue, electrode arrays record from a small fraction of cells in a given volume, limiting the validity of inference of the pairwise conditional dependencies. This may be the reason that partial correlations have not previously been used to describe  functional connectivity in cortical populations.

Two-photon imaging of population calcium signals presents unique advantages for the estimation of functional connectivity.  While the temporal resolution of calcium signals is limited by calcium dye kinetics, fast imaging techniques combined with spike inference algorithms provide millisecond-scale temporal resolution of single action potentials \cite{Grewe:2010}. However, such high temporal precision comes at the cost of the accuracy of inferred spike rates.  Better accuracy is achieved when calcium signals are analyzed on scales of tens of milliseconds \cite{Cotton:2013}.  The major advantage of calcium imaging is its ability to characterize the spatial arrangement and types of recorded cells.  Recently, advanced imaging techniques have allowed for recordings from nearly every cell in a volume of cortical tissue  \emph{in vivo} \cite{Katona:2012, Cotton:2013} and even from entire nervous systems \cite{Leung:2013, Ahrens:2013}.  These techniques may provide more incisive measurements of functional connectivity than classical electrophysiological recordings.

@@

Coactivation patterns and pairwise connectivity are not mutually exclusive since assemblies arise from patterns of synaptic connectivity.  However, an analysis of coactivation shifts the focus from detailed interactions to  collective behavior.  
In our study, the functional connectivity through modes of coactivations was represented by the factor analysis estimator $C_{\sf factor}$.

@@

\paragraph{Addressing ill-posedness}
Inferring the conditional dependencies between variables in a probabilistic model is an ill-posed problem: small variations in the data produce large errors in the inferred network of dependencies.  \Kcomment{References for the previous statement?} The problem becomes worse as the number of  recorded neurons increases until such models lose their statistical validity \cite{Roudi:2009}.  As  techniques improved to allow for recording from larger neuronal populations, experimental neuroscientists addressed this problem by extending the recording durations to keep sampling noise in check and verified that existing models are not overfitted and that the stationarity assumption is observed\cite{Tkacik:2013}. \Kcomment{The stationarity assumption seems a bit orthogonal to the rest discussed here.}  However,  ambitious projects such as the BRAIN initiative  \cite{Alivisatos:2013}, aim to record from significantly larger populations. Simply increasing recording duration will not be sufficient, and the problem must be addressed using regularized estimators. Regularization biases the solution toward a small subspace to counteract the effect of  sampling noise in the empirical data. However, biasing the solution to an inappropriate subspace does not allow significant estimation improvement and hampers interpretation.

@@

The results of model selection must be interpreted with caution.  As we demonstrated by simulation, with limited data an incorrect, but simpler (lower dimension) model can produce a better estimate than a more complex model with correct representation of dependencies.   Therefore, showing that a more constrained model has better cross-validated performance than a more complex model does not support the conclusion that it reveals a better representation of dependencies in the data.  This caveat is related to \emph{Stein's Paradox} \cite{Efron:1977}: The biasing of an estimate toward an arbitrary target is guaranteed to produce \emph{some} improvement over the best unbiased estimate.

@@

Moreover, this estimation approach provides a graphical representation of the dependencies in the data that can be used to formulate and test hypotheses about the structure of connectivity in the microcircuit. Importantly, the inferred functional interactions were substantially different from those measured by considering the network of the most significant sample correlations \figref{05}{F, H, and I}.  For example, the $C_{\sf sparse+latent}$ estimator reveals a large number of negative interactions that were not present in the sample correlation matrix (\figref{05}{F}) that may reflect inhibitory circuitry.  Although the network of functional interactions inferred by the optimal estimator, $C_{\sf sparse+latent}$, will not be identical to the anatomical connectivity it may accurately reflect average connectivity profiles. Future experiments utilizing molecular markers for different cell types, and follow-up multiple whole-cell in vitro experiments \cite{Hofer:2012, Ko:2013} \Kcomment{ref} will directly compare the inferred functional connectivity graphs to the underlying anatomical circuitry. Finally, the latent units inferred by the estimator can be analyzed for their physiological interpretation. For example, these latent units may be modulated under different brain states (e.g. slow-wave sleep, attention) and stimulus conditions (e.g. certain types of stimuli may engage feedback connections)
